---
title: "Cleaning up the data set"
author: "Natasha Djuric"
date: "November 23, 2019"
output:
  html_document:
    df_print: paged
  pdf_document: default
---


###Set Up
Setting working directory and loading packages and data frames
```{r}

# 1. SETTING WORKING DIRECTORY
setwd("D:/Third Year/EEB313/Project")

# 2. LOADING PACKAGES
library(tidyverse) 
library(dplyr)

# 3. LOADING FILES

#MAIN, RAW DATA SET containing compilation from different studies
fishraw <- read.delim("Predator_and_prey_body_sizes_in_marine_food_webs_vsn4.txt") 

#SUMMARY OF CHARACTERISTICS FROM EACH PAPER (paper name, paper age, sample size, location)
summary <- read_csv("Paper_Summary.csv")

```

###Investigating the fish data
###MAIN CHANGES:

-Every data frame name (fish_filtered_1, fish_filtered_2...) will correspond to one code chunk 
in which a specific type of cleaning was done.

1. Remove unwanted variables (notably any standardized variable columns and the associated 
columns listing strategies with standardizing them)

2. Keep ectothermic vertebrate data and remove cephalopods because there wasn't a large 
enough and diverse enough sample size of cephalopods to make a comparison between the groups.

3. Rename incongruous level names (due to typos, spelling differences) of the 
Specific.habitat factor that actually represent the same type of habitat

4. Keeping all years, even though they cover a wide temporal range of 50 years, 
unless we find a specific study to be inexplicably anomalous, because we don't want to throw out data unnecessarily.

5. Convert all measurements into a single unit type and remove the units column 
(all lengths in cm, all masses in g for easier analysis)

6. Sort data by quality, specifically removing any "Unsuitable" measurements 
with a quality rating of 5.

7. Converted Lat/Long data into degree format to make mapping and converting 
into a csv easier.
```{r}
#Many unnecessary/vague variables.
colnames(fishraw)
```


```{r}

# There are no NAs according to this output. However, upon visual inspection 
#it seems like some actually exist but have been labelled as "n/a" 
#so they aren't showing up. It looks like the n/a are only in columns we aren't keeping anyway, so we'll
# worry about them if they appear in the near-complete data frame.
any(is.na(fishraw))

```

###Selecting columns of interest
```{r}
#Selecting all the columns we want to keep and saving them into a new object:

#Removed all standardized columns because different papers 
#performed standardization in different ways,
#and their exact methods were hard to determine.
#We prefer to make our own standardized column,
#if necessary, from the raw measurements.

fish_filtered_1 <- fishraw %>% 
  dplyr::select(c(Record.number, Individual.ID, Predator, Predator.common.name, 
                  Predator..taxon, 
           Predator.lifestage, Type.of.feeding.interaction,
           Predator.length, Predator.length.unit, Predator.dimension.measured,
           Predator.fork.length, Predator.total.length,
           Predator.quality.of.length.mass.conversion, Predator.mass, 
           Predator.mass.unit, Prey.length, Prey.length.unit, Prey.mass, 
           Prey.mass.unit, Geographic.location, Latitude, Longitude, Depth,
           Mean.annual.temp, Depth, Mean.PP, SD.PP, SD.annual.temp, Reference, 
           Specific.habitat, Notes...assumptions,Prey.quality.of.conversion.to.length, 
           Prey.quality.of.conversion.to.mass, Predator.quality.of.length.mass.conversion, 
           Prey.dimension.measured, Prey.common.name, Prey.taxon, Prey))


# Most of the predators are ectothermic vertebrates. We decided to remove 
# cephalopod observations and focus on ectothermic vertebrates, especially since 
# most of the cephalopods came from one paper and it would be hard to tease apart
# specific environmental effects and sampling method from the results, and there
# are no other cephalopods to compare to. 

fish_filtered_1<-fish_filtered_1 %>%
  filter(Predator..taxon != "cephalopod") 

```


###Unit conversions

-Converting all predator and prey lengths to cm
-Converting all prey mass to g.
-All predator mass in g already.
-Once finished, removed the unit columns.

```{r}
fish_filtered_2<-fish_filtered_1 %>% 
  mutate(Predator.length = case_when(
    Predator.length.unit == "mm" ~ Predator.length / 10,
    Predator.length.unit == "µm" ~ Predator.length / 10000,
    TRUE ~ Predator.length 
  )) %>% 
  mutate(Prey.mass = case_when(
    Prey.mass.unit == "mg" ~ Prey.mass / 1000,
    TRUE ~ Prey.mass
  )) %>% 
  mutate(Prey.length=case_when(
    Prey.length.unit=="mm"~Prey.length/10,
    Prey.length.unit=="µm"~Prey.length/10000,
    TRUE~Prey.length
  )) %>% 
  select(c(-Prey.mass.unit, -Predator.length.unit, -Predator.mass.unit, -Prey.length.unit))

```

###Renaming repetitive levels of factors

1. Habitats, feeding interactions, predator lifestage

We gave common habitats the same level name if they had slightly 
different spelling, capitalization, or wording.

```{r}
#Renaming similar habitats, capitalizing feeding interactions and predator
#lifestages for consistency

fish_filtered_3 <-fish_filtered_2 %>% 
  mutate(Specific.habitat=recode(Specific.habitat, "Coastal bay"="Coastal Bay", 
                                 "Nearshore waters" = "Nearshore", "nearshore"="Nearshore", 
                                 "shelf"="Shelf", "inshore"="Inshore", 
                                 "transition region"="Transition",
                                 "shelfbreak/open ocean"="Shelfbreak/Open Ocean", 
                                 "estuary/coastal"="Estuary/Coastal", 
                                 "offshelf and on shelf"="Off/On Shelf",
                                 "open ocean"="Open Ocean", 
                                 "demersal food web"="Sea Floor")) %>% 
  mutate(Predator.lifestage=recode(Predator.lifestage, "adult"="Adult",
                                   "juvenile"="Juvenile", "larva"="Larva")) %>% 
  mutate(Type.of.feeding.interaction = recode(Type.of.feeding.interaction,
                                              "planktivorous" = "Planktivorous", 
                                              "piscivorous" = "Piscivorous", 
                                              "predacious"="Predacious"))

```

###Removing data with low quality

We will also consider the conversion quality of the columns
Predator_quality_of_length_mass_conversion, Prey_quality_of_conversion_to_length, 
Prey_quality_of_conversion_to_mass, and remove data with a quality rating of 5
(quality is ranked on a scale: 0 = mass measured, 1 = species regression, 2 = genus regression, 
3 = family regression, 4 = general shape, 5 = unsatisfactory).

We kept 0-4 so as not to lose too much data, and we were satisfied 
with the different levels as long as we later account for them in analysis.

```{r}

fish_filtered_4<-fish_filtered_3 %>% 
  filter(Prey.quality.of.conversion.to.length %in% 0:4,
         Prey.quality.of.conversion.to.mass %in% 0:4, 
         Predator.quality.of.length.mass.conversion %in% 0:4) 
  
```


### Converting NSEW latitude/longitude coordinates into degrees 

-When the data gets exported to a csv, NSEW coordinates do not render properly.
-We convert them to degree format instead.
-Our initial use of calculations to convert NSEW to degrees 
did not give accurate conversions, so we
manually inputted them into Google Earth to be safe and included those values below.


```{r}
####Converting Coordinates to Decimal Degrees 

# convert latitude and longitude to character vectors 
fish_filtered_5 <- fish_filtered_4 %>% 
  mutate(Latitude = as.character(Latitude), Longitude = as.character(Longitude))

### Conversion of Longitude and Latitude to Decimal Degrees ###

# Apalachicola Bay, Florida
fish_filtered_5$Latitude[fish_filtered_5$Latitude == "29º40'N"] <- 29.67
fish_filtered_5$Longitude[fish_filtered_5$Longitude == "85º10'W"] <- -85.17

# Gulf of Alaska
fish_filtered_5$Latitude[fish_filtered_5$Latitude == "56º50'N"] <- 56.83
fish_filtered_5$Longitude[fish_filtered_5$Longitude == "156º00'W"] <- -156.00

# Jeffreys Ledge, Gulf of Maine, New England
fish_filtered_5$Latitude[fish_filtered_5$Latitude == "42º80'N"] <- 43.33
fish_filtered_5$Longitude[fish_filtered_5$Longitude == "70º10'W"] <- -70.17

# Stellwagen Bank, Gulf of Maine, New England
fish_filtered_5$Latitude[fish_filtered_5$Latitude == "42º40'N"] <- 42.67
fish_filtered_5$Longitude[fish_filtered_5$Longitude == "70º15'W"] <- -70.25

# Cape Cod Bay, Gulf of Maine, New England
fish_filtered_5$Latitude[fish_filtered_5$Latitude == "41º86'N"] <- 42.43
fish_filtered_5$Longitude[fish_filtered_5$Longitude == "70º42'W"] <- -70.70

# Great Channel South, Gulf of Maine, New England
fish_filtered_5$Latitude[fish_filtered_5$Latitude == "41º71'N"] <- 42.18
fish_filtered_5$Longitude[fish_filtered_5$Longitude == "69º76'W"] <- -69.76

# South of Marthas Vineyard, New England
fish_filtered_5$Latitude[fish_filtered_5$Latitude == "41º08'N"] <- 41.13
fish_filtered_5$Longitude[fish_filtered_5$Longitude == "71º37'W"] <- -71.62

# Gulf of Maine, New England
fish_filtered_5$Latitude[fish_filtered_5$Latitude == "42º79'N"] <- 43.32
fish_filtered_5$Longitude[fish_filtered_5$Longitude == "69º51'W"] <- -69.85

# Andaman Sea (West of South Thailand)
fish_filtered_5$Latitude[fish_filtered_5$Latitude == "08º24'N"] <- 8.40
fish_filtered_5$Longitude[fish_filtered_5$Longitude == "97º53'E"] <- 97.88

# NE North Sea and Skaggerak
fish_filtered_5$Latitude[fish_filtered_5$Latitude == "57º00'N"] <- 57.00
fish_filtered_5$Longitude[fish_filtered_5$Longitude == "08º00'E"] <- 8.00

# W Greenland
fish_filtered_5$Latitude[fish_filtered_5$Latitude == "66º20'N"] <- 66.33
fish_filtered_5$Longitude[fish_filtered_5$Longitude == "56º00'W"] <- -56.00

# French Polynesian EEZ
fish_filtered_5$Latitude[fish_filtered_5$Latitude == "12º00'S"] <- -12.00
fish_filtered_5$Longitude[fish_filtered_5$Longitude == "144º00'W"] <- -144.00

# Catalan Sea
fish_filtered_5$Latitude[fish_filtered_5$Latitude == "40º55'N"] <- 40.92
fish_filtered_5$Longitude[fish_filtered_5$Longitude == "02º40'E"] <- 2.67

# Western North Pacific
fish_filtered_5$Latitude[fish_filtered_5$Latitude == "37º00'N"] <- 37.00
fish_filtered_5$Longitude[fish_filtered_5$Longitude == "143º00'E"] <- 143.00

# Atlantic Ocean
fish_filtered_5$Latitude[fish_filtered_5$Latitude == "40º00'N"] <- 40.00
fish_filtered_5$Longitude[fish_filtered_5$Longitude == "71º00'W"] <- -71.00

# Off the Bay of Biscay (Depth = 4562)
fish_filtered_5$Latitude[fish_filtered_5$Latitude == "45º00'N"] <- 45.00
fish_filtered_5$Longitude[fish_filtered_5$Longitude == "18º00'W"] <- -18.00

# Off the Bay of Biscay (Depth = 3798)
fish_filtered_5$Latitude[fish_filtered_5$Latitude == "44º00'N"] <- 44.00
fish_filtered_5$Longitude[fish_filtered_5$Longitude == "16º00'W"] <- -16.00

# Great South Bay, Long Island, New York
fish_filtered_5$Latitude[fish_filtered_5$Latitude == "40º10'N"] <- 40.17
fish_filtered_5$Longitude[fish_filtered_5$Longitude == "73º10'W"] <- -73.17

# Strait of Georgia
fish_filtered_5$Latitude[fish_filtered_5$Latitude == "49º00'N"] <- 49.00
fish_filtered_5$Longitude[fish_filtered_5$Longitude == "123º00'W"] <- -123.00

# Europe, Celtic Sea ecosystem
fish_filtered_5$Latitude[fish_filtered_5$Latitude == "50º50'N"] <- 50.83
fish_filtered_5$Longitude[fish_filtered_5$Longitude == "08º00'W"] <- -8.00

# Antarctic Peninsula (Depth = 881)
fish_filtered_5$Latitude[fish_filtered_5$Latitude == "63º00'S"] <- -63.00
fish_filtered_5$Longitude[fish_filtered_5$Longitude == "58º00'W"] <- -58.00

# Antarctic Peninsula (Depth = 1752)
fish_filtered_5$Latitude[fish_filtered_5$Latitude == "62º00'S"] <- -62.00
fish_filtered_5$Longitude[fish_filtered_5$Longitude == "55º00'W"] <- -55.00

# Oxwich Bay, Wales
fish_filtered_5$Latitude[fish_filtered_5$Latitude == "51º52'N"] <- 51.87
fish_filtered_5$Longitude[fish_filtered_5$Longitude == "04º10'W"] <- -4.17

# Greenland 
fish_filtered_5$Latitude[fish_filtered_5$Latitude == "60º00'N"] <- 60.00
fish_filtered_5$Longitude[fish_filtered_5$Longitude == "40º00'W"] <- -40.00

# Eastern Mediterranean
fish_filtered_5$Latitude[fish_filtered_5$Latitude == "38º00'N"] <- 38.00
fish_filtered_5$Longitude[fish_filtered_5$Longitude == "23º00'E"] <- 23.00


```


###Mass ratio calculation and log conversion
```{r}
fish_filtered_6<-fish_filtered_5 %>% 
  mutate(massratio=(Predator.mass/Prey.mass)) %>% 
  mutate(massratiolog=log10(massratio)) 
```

###Exporting cleaned up data to a csv
```{r}
#dir.create("data-processed")
#write_csv(fish_filtered_6, "data-processed/fish_clean_final.csv")

```
