---
title: "Linear Mixed Effect Model Selection and Plotting the Results"
author: "Natasha Djuric"
output:
  html_document:
    df_print: paged
---

###LINEAR MIXED EFFECTS MODEL SELECTION

-We're going to avoid including factors with too many levels as fixed effects
-Make a series of models with diff fixed and random effects, use AIC to distinguish them

```{r}
# SET UP 

#Loading packages
library(tidyverse)
library(lme4)
library(lmerTest)
library(MuMIn)

#Cleaned up data
fish<-read_csv("fish_clean_final.csv")

```

Prey quality may influence our interpretation of prey mass effects on log predator mass. 
When we do an ANOVA, we see that different prey qualities have  different log predator masses. 
To maintain statistical power, we will use all data with a prey quality of 4 
because it has the highest sample size and we are comfortable with using "general shape" 
for prey mass, in order to predict predictor mass.

```{r}
#We're only going to focus on a prey quality of 4 (general shape)
fish_filter <- fish %>%
 filter(Prey.quality.of.conversion.to.mass == "4" ) 

#Note for model analysis: There are no predator qualities of 
#0 that also have a prey quality of 4
table(fish$Prey.quality.of.conversion.to.mass, 
      fish$Predator.quality.of.length.mass.conversion)

```

###STANDARDIZING 

"It is good practice to standardise your explanatory variables 
before proceeding so that they have a mean of zero (“centering”) 
and standard deviation of one (“scaling”). It ensures that 
the estimated coefficients are all on the same scale, 
making it easier to compare effect sizes" 
- from https://ourcodingclub.github.io/2017/03/15/mixed-models.html. 

For our Mean PP (Primary Productivity) variable, there is only 
one value for each geographic location. The authors provided the standard deviation for 
each mean value in a second column. 

```{r}
#Standardizing mean PP by dividing by the standard deviation
fish_filter<-fish_filter %>% 
  mutate(Mean_PP_Standardized= Mean.PP/SD.PP)


#Standardizing prey mass
fish_filter<-fish_filter %>% 
  mutate(Prey.mass.standardized = scale(Prey.mass, center=TRUE, scale=TRUE))
```

###CHECKING ASSUMPTIONS OF LMEM

*Normality at each X and homogeneity of variances at each X:*
Not that important with our large sample size, 
but log10 is more normal so we will use it.

```{r}

hist(x=(fish_filter$Predator.mass))
hist(x=log10(fish_filter$Predator.mass))

#Converting predator mass to log mass for normality
fish_filter<-fish_filter %>% 
  mutate(logpredmass = log10(Predator.mass))
```
*Fixed X*

All our predictors are in the range of values we'd expect. Standardized prey mass 
has many negative values, so we can't do a log transformation, 
but as explanatory variables they 
do not need to be normally distributed regardless. 
From 
https://www.theanalysisfactor.com/the-distribution-of-independent-variables-in-regression-models/

```{r}
ggplot(fish_filter, aes(x=Mean_PP_Standardized))+geom_histogram()
ggplot(fish_filter, aes(x=Geographic.location, y=Mean_PP_Standardized))+geom_boxplot()

#Standardized prey mass
ggplot(fish_filter, aes(x=(Prey.mass.standardized)))+geom_histogram()

```

*Independence of observations:*
Some of the fish come from the same stomach. However, we don't need to nest the 
data by record ID because we're interested in predicting Predator mass, not Prey mass. 
We should, however, nest the predators within geographic location for several reasons: 
in some instances, predators were not a random sample from the geographic location. 
Also, each "Geographic.location" acts as a proxy for the year and sampling methods of a study,
and is our best way for accounting for spatial autocorrelation.

Therefore, going to add (1|Geographic.location) as a random effect to account for non-independence.

*Controling for quality*
```{r}

# #LMEMs sometimes have a hard time estimating random effect components 
# when the random effects have too few levels (i.e., < 5 or 6). 
# In fact, when we included Predator.quality as a random effect (4 levels),
# the model could not estimate the variance. Therefore, we will use 
#Predator quality as a fixed effect without interpreting its coefficients
#because we know different predator qualities have significantly different
# predator masses (ANOVA), also visible in histograms below.

#Converting quality to factor
fish_filter$Predator.quality.of.length.mass.conversion<-
  as.factor(fish_filter$Predator.quality.of.length.mass.conversion)

ggplot(fish_filter, aes(y=logpredmass, 
                        x=Predator.quality.of.length.mass.conversion))+geom_boxplot()



#Reducing levels of lifestage to remove ambiguity
fish_filter<-fish_filter %>% 
filter(Predator.lifestage !="larva / juvenile",
       Predator.lifestage != "postlarva", Predator.lifestage !="postlarva/juvenile") 

```
STEP 1: SATURATED MODEL

-Predator taxon cannot be a fixed effect. There are too many levels and it won't run.
-It is best to encode as random effect.
-We don't care about how specific predators are related to prey anyway -- 
all we care about are general trends. Given that we're asking about
the general relationship between size and feeding interaction, age, etc, 
species are really just a random effect anyway: we are pulling out general trends.

```{r}
mixed_model_saturated <- lmer(logpredmass ~ Mean_PP_Standardized + 
                                Predator.lifestage + Type.of.feeding.interaction + 
                                Prey.mass.standardized + 
                                Predator.lifestage:Type.of.feeding.interaction +
                                Predator.quality.of.length.mass.conversion +
                                (1|Predator.common.name)+(1|Geographic.location),  
                              data=fish_filter, REML=FALSE)
```


STEP 2: OPTIMIZING RANDOM EFFECTS STRUCTURE 

```{r}
# #Leave mass conversion random effect in all models because the 
# response variable was calculated using this factor.
# #Predator species has to stay in the model because the 
# researchers specifically 
# #sought these species, they aren't a random sample from
# that environment. 

#We have only non-independent random effects at this point: nothing we can remove.

mixed_model_saturated_True<-lmer(logpredmass ~ Mean_PP_Standardized + 
                                   Predator.lifestage + Type.of.feeding.interaction + 
                                   Prey.mass.standardized + 
                                   Predator.lifestage:Type.of.feeding.interaction + 
                                   Predator.quality.of.length.mass.conversion +  
                                   (1|Predator.common.name)+
                                   (1|Geographic.location),  
                                 data=fish_filter, REML=TRUE)


#AIC or AICc Calculation using the most saturated model
#n/k < 40 use AICc
#k=20
#AIC ratio is 1414.077
nrow(fish_filter)/13


#Therefore, we will use AIC. 
#The optimal (only) random effects structure is the saturated model.
```

STEP 3. CHOOSING FIXED EFFECTS 
We now need to refit the saturated model (which has the optimal 
random-effect structure with REML=true) as false and 
compare different fixed-effect structures.

```{r}
#Full model with all fixed effects and their interactions

# mixed_model_saturated <- lmer(logpredmass ~ Mean_PP_Standardized + 
# Predator.lifestage + Type.of.feeding.interaction + Prey.mass.standardized + 
#   Predator.lifestage:Type.of.feeding.interaction + 
#   Predator.quality.of.length.mass.conversion + (1|Predator.common.name)+
#   (1|Geographic.location),  data=fish_filter, REML=FALSE)

 
#INTERACTIONS + SOME FIXED EFFECTS --------------------------------------------------------

#No Mean_PP_Standardized
mixed_model_interactions_no_PP<-lmer(logpredmass ~ Predator.lifestage
                                     + Type.of.feeding.interaction +
                                       Prey.mass.standardized + 
                                       Predator.lifestage:Type.of.feeding.interaction + 
                                       Predator.quality.of.length.mass.conversion + 
                                       (1|Predator.common.name)+
                                       (1|Geographic.location), 
                                     data=fish_filter, REML=FALSE)


#No Prey.mass.standardized 
mixed_model_interactions_no_prey_mass<-
  lmer(logpredmass ~ Mean_PP_Standardized + 
Predator.lifestage + Type.of.feeding.interaction + 
  Predator.lifestage:Type.of.feeding.interaction + 
  Predator.quality.of.length.mass.conversion + (1|Prey.taxon) +
  (1|Predator.common.name)+(1|Geographic.location),  data=fish_filter, REML=FALSE)


#COMBINATIONS OF NO INTERACTIONS -----------------------------------------------

#No interaction
mixed_model_no_interactions<-
  lmer(logpredmass ~ Mean_PP_Standardized + Predator.lifestage + 
         Type.of.feeding.interaction + Prey.mass.standardized +  
         Predator.quality.of.length.mass.conversion + 
         (1|Predator.common.name)+(1|Geographic.location),  
       data=fish_filter, REML=FALSE)

#No interactions or Mean_PP
mixed_model_no_interactions_no_PP<-
  lmer(logpredmass ~ Predator.lifestage + Type.of.feeding.interaction + 
         Prey.mass.standardized +  
         Predator.quality.of.length.mass.conversion + 
         (1|Predator.common.name)+(1|Geographic.location),  data=fish_filter, REML=FALSE)

#No interactions or Predator.lifestage
mixed_model_no_interactions_no_lifestage<-
  lmer(logpredmass ~ Mean_PP_Standardized + Type.of.feeding.interaction + 
         Prey.mass.standardized + 
         Predator.quality.of.length.mass.conversion + 
         (1|Predator.common.name)+(1|Geographic.location),  
       data=fish_filter, REML=FALSE)

#No interactions or type of feeding interaction
mixed_model_no_interactions_no_feeding<-
  lmer(logpredmass ~ Mean_PP_Standardized + Predator.lifestage + 
         Prey.mass.standardized + Predator.quality.of.length.mass.conversion +
         (1|Predator.common.name)+(1|Geographic.location),  data=fish_filter, REML=FALSE)

#No interactions or Prey.mass.standardized
mixed_model_no_interactions_no_preymass<-
  lmer(logpredmass ~ Mean_PP_Standardized + Predator.lifestage + 
         Type.of.feeding.interaction +
         Predator.quality.of.length.mass.conversion  + 
         (1|Predator.common.name)+(1|Geographic.location), 
       data=fish_filter, REML=FALSE)

#Testing fixed-effects models
AIC(mixed_model_saturated, mixed_model_interactions_no_PP,
    mixed_model_interactions_no_prey_mass, mixed_model_no_interactions,
    mixed_model_no_interactions_no_PP, 
    mixed_model_no_interactions_no_lifestage, 
    mixed_model_no_interactions_no_feeding, 
    mixed_model_no_interactions_no_preymass)

#The top two models are 1) mixed_model_interactions_no_PP and 2) the saturated model.
#We considered averaging the models, but given that there was a minimal difference in them
#except for the presence of PP, which was not even a significant predictor in the model
#that was present and upon graphing no trend was visible, we just stuck with
#the mixed_model_interactions_no_PP model.
summary(mixed_model_interactions_no_PP)
summary(mixed_model_saturated)


# load packages for html table model output
library(sjPlot)
library(sjmisc)
library(sjlabelled)

#Outputs a cleaned up version of the model.
#tab_model(mixed_model_interactions_no_PP)

```

###PLOTTING

```{r}
#Predator mass by lifestage: as they get bigger, they get older
ggplot(fish_filter, aes(x=Predator.lifestage, y=logpredmass))+
  geom_boxplot()+
  theme_bw()+scale_x_discrete(limits=c("Larva","Juvenile","Adult"))+
  labs(y="Log (Predator mass)", x="Predator Lifestage")
                                                           
#ggsave("Lifestage_LogPredMass.jpg", plot = last_plot(), device = NULL, path = NULL,
#  scale = 1, width = NA, height = NA, units = c("in", "cm", "mm"),
#  dpi = 300, limitsize = TRUE)

```

PLOTTING PRIMARY PRODUCTIVITY
```{r}
#Primary productivity is removed from the model. This is not unreasonable, given that there
#is no obvious relationship when you plot the points: 
#We do see that the shelf habitat shows considerable variation in 
#predator mass for same mean primary productivity, 
#suggesting a greater diversity in that environment. This relationship just isn't linear.

library(scales)

ggplot(fish_filter, aes(x=Mean_PP_Standardized, y=log10(Predator.mass), 
                        colour=Specific.habitat))+
  geom_boxplot(width=0.8)+theme_bw()+
  labs(x="Mean Primary Productivity (Standardized)", 
       y="log (Predator mass)", fill="Geographic Location")+
  scale_x_continuous(labels=comma)

#ggsave("MPP_LogPredMass.jpg", plot = last_plot(), device = NULL, path = NULL,
#  scale = 1, width = NA, height = NA, units = c("in", "cm", "mm"),
#  dpi = 300, limitsize = TRUE)

```
PLOTTING PREY MASS

```{r}
#Prey mass significantly and positively predicts predator mass, but weakly.
#This is reflected by the low coefficient of our model
#Trend is primarily driven by Atlantic ocean fish
#Removed outliers to make trend more obvious
#Most slopes are positive

ggplot(fish_filter, aes(x=Prey.mass.standardized,
                        y=logpredmass, colour=Geographic.location))+
  geom_point(alpha=0.3)+xlim(c(0,34))+ylim(0,6)+
  theme_classic()+
  labs(x="Prey mass (standardized)", y="Predator mass (log)")+
  geom_smooth(method="lm", se=F)

# ggsave("Prey_LogPredMass.jpg", plot = last_plot(), device = NULL, path = NULL,
#   scale = 1, width = NA, height = NA, units = c("in", "cm", "mm"),
#  dpi = 300, limitsize = TRUE)

```

PLOTTING FEEDING INTERACTION 

```{r}

 ggplot(fish_filter,aes(x=Type.of.feeding.interaction, y=logpredmass)) +
  geom_boxplot()+
  theme_classic() +labs(y="Log (Predator mass)",
                        x="Predator Feeding Interaction")+
                        scale_x_discrete(limits=c("Planktivorous","Piscivorous","Predacious"))         
                                           

# ggsave("FeedingInteraction_LogPredMass.jpg", plot = last_plot(), 
#device = NULL, path = NULL,
#   scale = 1, width = NA, height = NA, units = c("in", "cm", "mm"),
#   dpi = 300, limitsize = TRUE)


```
PLOTTING THE INTERACTION BETWEEN LIFESTAGE AND FEEDING INTERACTION 

```{r}
fish_filter %>%
filter(Predator.lifestage != "Larva", Type.of.feeding.interaction != "planktivorous") %>%
group_by(Predator.lifestage, Type.of.feeding.interaction) %>%
summarise(meanlogpredmass = mean(logpredmass), n_obs = n(),
se_mass = (meanlogpredmass/sqrt(n_obs))) %>%
ggplot()+
aes(x = Predator.lifestage, y = meanlogpredmass, colour = Type.of.feeding.interaction)+
geom_line(aes(group=Type.of.feeding.interaction))+
geom_errorbar(aes(ymin = (meanlogpredmass-se_mass), ymax = (meanlogpredmass + se_mass)),
width = 0.05)+
geom_point()+
labs(x = "Predator Lifestage", y = "Mean of Log Predator Mass",
colour = "Type of Feeding Interaction")+
scale_x_discrete(limits = c("Juvenile", "Adult"))+
scale_colour_manual(values = c("midnightblue", "orangered"))+
theme_classic()
```
